---
title: "Example for spSeudoMap in R (real sorted data)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spSeudoMap_sorted}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
)
```

## Install required packages
```{r setup}
# ## Install devtools
# if (!requireNamespace("devtools", quietly = TRUE))
#   install.packages('devtools')
# 
# ## Install spSeudoMap package
# if (!requireNamespace("spSeudoMap", quietly = TRUE))
#   devtools::install_github("bsungwoo/spSeudoMap")
#
```

## Load required packages
```{r}
# library(spSeudoMap)
# library(SeuratObject)
```


## Make output and data directories
```{r}
# # Set the location of the R project file as a default directory
# filepath <- rprojroot::find_rstudio_root_file()
# # Set working directoy
# setwd(filepath)
# 
# # Create output directory
# output_folder_name <- 'output'
# if (!file.exists(output_folder_name)){
#   dir.create(output_folder_name)
# }
# 
# # Create data directory
# if (!file.exists('data')){
#   dir.create('data')
# }
```


## Install conda environment
<!-- conda create -n spSeudoMap python=3.7.12 -c conda-forge -->
<!-- conda activate spSeudoMap -->
<!-- pip install git+https://github.com/bsungwoo/spSeudoMap.git -->
<!-- python -m ipykernel install --user --name spSeudoMap --display-name spSeudoMap -->


## Load single-cell and spatial datasets
### Load single-nucleus data and generate simulation dataset
```{r}
# # Seurat object file for single-cell data stored in sc_data.rds
# sc_data <- readRDS('sc_data.rds')
```


### Load spatial dataset (RDS file with Seurat object): 10X genomics data repository
#### V1_Adult_Mouse_Brain_Coronal_Section_1
```{r}
# # Seurat object file for spatial data stored in sc_data.rds
# sp_data <- readRDS('sp_data.rds')
```

### Check the size of spatial data
```{r}
# dim(sp_data)
```

### Set the number of pseudospots: 5 times the number of spatial spots
```{r}
# npseudo <- 5*dim(sp_data)[2]
# npseudo
```

### Set the pseudotype fraction from the literature evidence or cell sorting experiment (fraction of non-sorted cell population)
```{r}
# # Presumed average pseudotype (non-sorted cell population) fraction: from the literature or cell sorting experiment
# pseudo_frac_m <- 0.5
```

## Perform spSeuoMap analysis
#### Explanation for the function
```{r}
# help(pred_cellf_spSeudoMap)
```

### Run spSeudoMap
```{r}
# # Using already installed conda environment
# # Conda environment will be automatically installed if 'spSeudoMap' is not already installed
# sp_data_cellf <- pred_cellf_spSeudoMap(sp_data,sc_data,
#                                        outdir=output_folder_name,
#                                        sp_subset=F,spot.cluster.name='seurat_clusters',
#                                        spot.cluster.of.interest=NULL,
#                                        env.select='conda',python.install=T,
#                                        python_path=NULL,env.name='spSeudoMap',
#                                        gpu=TRUE,metadata_celltype='annotation_1',
#                                        num_markers=40,mixture_mode='pseudotype',
#                                        seed_num=0,
#                                        mk_ratio_fix=T, mk_ratio=4,
#                                        pseudo_frac_m=pseudo_frac_m, pseudo_frac_std=0.1,
#                                        nmix=8,npseudo=npseudo,alpha=0.6,alpha_lr=5,
#                                        emb_dim=64,batch_size=512,n_iterations=3000,init_train_epoch=10)

```


### Save seurat object with cell fraction
```{r}
# saveRDS(sp_data_cellf, file.path(filepath, output_folder_name, 'sp_data_cellf.rds'))
```



### Visualization of spatial cell fraction
#### Remove '_cellf' from the column names cell fraction metadata
```{r}
# cellf.data <- sp_data_cellf@meta.data
# cellf.data.colname <- sapply(colnames(cellf.data), function(x){
#   if (grepl('_cellf',x)){return(strsplit(x, split='_cellf')[[1]][1])}
#   else {return(x)}
# })
# sp_data_cellf.mod <- sp_data_cellf
# colnames(sp_data_cellf.mod@meta.data) <- cellf.data.colname
```


#### Visualize the cell types
```{r}
# p <- Seurat::SpatialFeaturePlot(sp_data_cellf.mod, features = cellf.data.colname[1:8], 
#                                 ncol = 4, alpha=0.6, combine = FALSE)
# patchwork::wrap_plots(p, ncol = 4)

```
