---
title: "Example for spSeudoMap in R (Simulation data)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
)
```

## Install required packages
```{r setup}
## Install devtools
if (!requireNamespace("devtools", quietly = TRUE))
  install.packages('devtools')

## Install spSeudoMap package
if (!requireNamespace("spSeudoMap", quietly = TRUE))
  devtools::install_github("bsungwoo/spSeudoMap")

## Install sceasy
# https://github.com/cellgeni/sceasy
# 1. Install BiocManager
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

# 2. Install LoomExperiment
if (!requireNamespace("LoomExperiment", quietly = TRUE))
  BiocManager::install(c("LoomExperiment", "SingleCellExperiment"))

# 3. Install sceasy
if (!requireNamespace("sceasy", quietly = TRUE))
  devtools::install_github("cellgeni/sceasy")

```

## Load required packages
```{r}
library(spSeudoMap)
library(SeuratObject)
```


## Make output and data directories
```{r}
# Set the location of the R project file as a default directory
filepath <- rprojroot::find_rstudio_root_file()
# Set working directoy
setwd(filepath)

# Create output directory
output_folder_name <- 'output'
if (!file.exists(output_folder_name)){
  dir.create(output_folder_name)
}

# Create data directory
if (!file.exists('data')){
  dir.create('data')
}
```


## Install conda environment
<!-- conda create -n spSeudoMap python=3.7.12 -c conda-forge -->
<!-- conda activate spSeudoMap -->
<!-- pip install git+https://github.com/bsungwoo/spSeudoMap.git -->
<!-- python -m ipykernel install --user --name spSeudoMap --display-name spSeudoMap -->


## Load single-cell and spatial datasets
### Download single-nucleus dataset
#### Kleshchevnikov, V., Shmatko, A., Dann, E. et al. Cell2location maps fine-grained cell types in spatial transcriptomics. Nat Biotechnol (2022).
#### https://cell2location.readthedocs.io/en/latest/notebooks/cell2location_short_demo.html¶
```{r}
options(timeout=1000)
download.file(url = "https://cell2location.cog.sanger.ac.uk/tutorial/mouse_brain_snrna/regression_model/RegressionGeneBackgroundCoverageTorch_65covariates_40532cells_12819genes/sc.h5ad",
              destfile = './data/adata_single.h5ad')
```


### Load single-nucleus data and generate simulation dataset
#### Load conda environment and load .h5ad
```{r}
## Load conda environment and convert to Seurat object
reticulate::use_condaenv("spSeudoMap", required=TRUE)
sc <- reticulate::import('scanpy', convert = FALSE)
sp <- reticulate::import('scipy.sparse', convert = FALSE)

# Load single-cell data 
sc_data <- sc$read(file.path(filepath, './data/adata_single.h5ad'))

```

#### Extract raw data, convert to the matrix and extract metadata
```{r}
# Extract raw data
sc_data = sc_data$raw$to_adata()

# Convert scipy matrix to sparse matrix in R
X <- Matrix::t(reticulate::py_to_r(sp$csc_matrix(sc_data$X)))
# Assign row names and column names
rownames(X) <- reticulate::py_to_r(sc_data$var['SYMBOL']$tolist())
colnames(X) <- reticulate::py_to_r(sc_data$obs$index$tolist())

# Convert metadata(.obs) in pandas dataframe into R dataframe
obs <- reticulate::py_to_r(sc_data$obs)
```

#### Create Seurat Object and save
```{r}
# Convert to Seurat object and add metadata
sc_data_r <- Seurat::CreateSeuratObject(X)
sc_data_r <- Seurat::AddMetaData(sc_data_r, obs)
```

#### (Load the Seurat object) and create simulation subpopulation data (with Ext neurons)
```{r}
# Subset excitatory neurons
ext_neuron_types <- grep(pattern='^Ext_', 
                         levels(factor(sc_data_r[['annotation_1']][,1])), value=T)
ext_neuron_types

# Set the identity to the cell types
Seurat::Idents(sc_data_r) <- sc_data_r[['annotation_1']]

# Subset the single-cell data to contain only the excitatory neurons
sc_data_sim <- subset(sc_data_r, idents=ext_neuron_types)

# Save the subsetted simulated single-cell data (Ext)
saveRDS(sc_data_sim, file.path(filepath,'./data/sc_data_sim.rds'))
```


### Load spatial dataset (RDS file with Seurat object): 10X genomics data repository
#### V1_Adult_Mouse_Brain_Coronal_Section_1
```{r}
# Import dataset from scanpy
sc$datasets$visium_sge(
  sample_id="V1_Adult_Mouse_Brain_Coronal_Section_1"

)

sp_data <- Seurat::Load10X_Spatial(file.path(filepath,'./data/V1_Adult_Mouse_Brain_Coronal_Section_1'))
```

### Check the size of spatial data
```{r}
dim(sp_data)
```

### Set the number of pseudospots: 5 times the number of spatial spots
```{r}
npseudo <- 5*dim(sp_data)[2]
npseudo
```

### Set the pseudotype fraction from the simulation dataset (fraction of non-excitatory neuron population)
```{r}
# Calculate pseudotype fraction from the simulation dataset
pseudo_frac_m <- 1 - ((dim(sc_data_sim)[2])/(dim(sc_data_r)[2]))
pseudo_frac_m
```

## Perform spSeuoMap analysis
### Cell type mapping of spatial transcriptomics using unmatched single-cell RNA-seq data

#### Seurat object to AnnData conversion algorithm: reference to sceasy
#### https://github.com/cellgeni/sceasy

#### Input
#### adata_sp: spatial data (AnnData object) to be used in predicting cell fraction (default: None)
#### -> count matrix should be non-normalied raw data
#### adata_sc: single-cell data (AnnData object) to be used in making pseudospots (default: None)
#### -> count matrix should be non-normalied raw data

#### outdir: the directory to save output files (models and results) (default = '.')
#### source.code.dir: directory of python source code (default = 'pred_cellf_spSeudoMap.py')

#### sp_subset: whether to subset spatial data and calculate for specific spot cluster (default = FALSE)
#### spot.cluster.name: group name of the cluster used for subsetting spatial data (default: 'seurat_clusters')
#### spot.cluster.of.interest: name of each spot clusters to be used (default: NULL)
#### metadata_celltype: column name for single-cell annotation data in metadata (default: 'celltype')

#### env.select: select between using reticulate virtual environment or conda environment ("virtual" or "conda")
#### python.install: whether to automatically install python version 3.7.12

#### python_path: path for the python 3.7. (default: NULL)
#### If NULL, python version 3.7.12 will be installed (valid for Linux) 
#### If "current", python interpreter associated with current virtual env (ex: r-reticulate) will be used. (version should be 3.7)
#### env.name: name of the virtual or conda environment to use for CellDART analysis (default: 'spSeudoMap')

#### gpu: check whether to use gpu (True) or not (False) (default = True)
#### metadata_celltype: column name for single-cell annotation data in metadata (default: 'celltype')
#### num_markers: number of selected marker genes in each cell-type (default = 20)

#### mixture_mode: mode of the pseudospot generation 
#### -> 'default': when cell types are similar between single-cell/spatial data (Identical to CellDART)
#### -> 'pseudotype': when there is a mismatch of cell types between single-cell/spatial data and cell types exclusively present in spatial data is considered when generating pseudospots

#### seed_num: seed to be used in random sampling (default = 0)

#### mk_ratio_fix: whether to fix the mk_ratio when selecting the 
#### mk_ratio: ratio of number of single-cell markers to virtual pseudotype markers (default = 2)
#### pseudo_num_genes: number of the virtual markers genes for pseudotype (default = 40)
### -> the number is used only when the mk_ratio is not fixed and number of pseudotype markers should be manually provided 

#### pseudo_frac_m: average of presumed fraction of the pseudotype (cell types exclusively present in spatial data) across all spatial spots (default = 0.5)
#### -> determined by cell sorting study or by literature based evidence that enables speculation of average fraction in the tissue
#### pseudo_frac_std: standard deviation of the distribution of presumed pseudotype fraction across all spatial spots (default = 0.1)
#### num_top_genes: number of top genes having highest log fold change between spatial and single-cell normalized pseudobulk counts (spatial/single-cell)
#### -> the top genes are used for generating module score to predict pseudotype fraction in spatial spots

#### nmix: the number of cells sampled from single-cell data when making a pseudospot (default = 10)
#### npseudo: a total number of pseudospots (default = 20000)

#### alpha: loss weights of domain classifier to the source classifier (default = 0.6)
#### alpha_lr: learning rate for the domain classifier (alpha_lr*0.001, default = 5)
#### emb_dim: output size of dimensions for feature extractor (default = 64)

#### batch_size: minibatch size for pseudospots and spatial data during the training (default = 512)
#### n_iterations: iteration number for the adversarial learning (default = 3000)
#### init_train_epoch: iteration number for the pre-training process (default = 10)

#### Output
#### sp_data_sub: spatial data (Seurat object) with predicted cell fraction in metadata (@meta.data)


### 1. Example for using conda environment (recommended)
#### For Linux distributions, conda environment will be automatically installed without previous installation
#### For Windows OS, install conda environment first and then call it using env.name='spSeudoMap'
#### For Windows OS, 'python.install' should be F
```{r}
# Using already installed conda environment
# Conda environment will be automatically installed if 'spSeudoMap' is not already installed
sp_data_cellf <- pred_cellf_spSeudoMap(sp_data,sc_data_sim,
                                       outdir=output_folder_name,
                                       sp_subset=F,spot.cluster.name='seurat_clusters',
                                       spot.cluster.of.interest=NULL,
                                       env.select='conda',python.install=T,
                                       python_path=NULL,env.name='spSeudoMap',
                                       gpu=TRUE,metadata_celltype='annotation_1',
                                       num_markers=40,mixture_mode='pseudotype',
                                       seed_num=0,
                                       mk_ratio_fix=T, mk_ratio=4,
                                       pseudo_frac_m=pseudo_frac_m, pseudo_frac_std=0.1, 
                                       nmix=8,npseudo=npseudo,alpha=0.6,alpha_lr=5,
                                       emb_dim=64,batch_size=512,n_iterations=3000,init_train_epoch=10)

```


### Save seurat object with cell fraction
```{r}
saveRDS(sp_data_cellf, file.path(filepath, output_folder_name, 'sp_data_cellf.rds'))
```



### Visualization of spatial cell fraction
#### Remove '_cellf' from the column names cell fraction metadata
```{r}
cellf.data <- sp_data_cellf@meta.data
cellf.data.colname <- sapply(colnames(cellf.data), function(x){
  if (grepl('_cellf',x)){return(strsplit(x, split='_cellf')[[1]][1])}
  else {return(x)}
})
sp_data_cellf.mod <- sp_data_cellf
colnames(sp_data_cellf.mod@meta.data) <- cellf.data.colname
```


#### Visualize the layer-specific excitatory neuons
```{r}
cell_types <- c('Ext_L25','Ext_L23','Ext_L56','Ext_L5_1',
                'Ext_L6B','Ext_Hpc_CA1','Ext_Hpc_CA3',
                'Ext_Hpc_DG1','Ext_Thal_1','Ext_Pir','Ext_Amy_1','Others')

p <- Seurat::SpatialFeaturePlot(sp_data_cellf.mod, features = cell_types, 
                                ncol = 4, alpha=0.6, combine = FALSE)
patchwork::wrap_plots(p, ncol = 4)

```
